How fluid-mechanical erosion creates anisotropic porous media

### 2022-06-22 docs/PhysicsPaper/StokesErosion.tex
NM

\begin{abstract}
Using a Cauchy integral formulation of the boundary integral equations, we simulate the erosion a porous medium comprised of up to 100 solid bodies embedded in a Stokes flow. The grains of the medium are resolved individually and erode under the action of surface shear stress. Through nonlinear feedback with the surrounding flow fields, microscopic changes in grain morphology give way to larger-scale features in the medium such as channelization. The Cauchy-integral formulation and associated quadrature formulas enable us to resolve dense configurations of nearly contacting bodies.
We observe substantial anisotropy to develop over the course of erosion; that is, the configurations that result from erosion generally permit flow in the longitudinal direction more easily than in the transverse direction by up to a factor of five. These results suggest that the erosion of solid material from groundwater flows may contribute to the well-known anisotropy of natural porous media.
%Tie it to alternate hypothesis in groundwater.
\end{abstract}
\maketitle

@@

	Figure \ref{fig1} shows an example simulation in which 80 circular bodies of randomized sizes and positions are immersed in a Stokes flow that moves from left to right. Color depicts the local speed of the flow intervening between bodies to give a sense of where the flow is strongest. Over time, individual bodies erode in response to the shear stresses induced on them.
The heterogeneous material removal creates visible features both at the level of individual bodies and the larger-scale configuration. For example, horizontally-oriented channels are clearly visible in the third and fourth frames.

%which regions of high flow speed create higher local shear rates, thus preferentially removing solid material, and further opening the channels. The positive feedback creates the channels that are clearly visible by in final two images.

@@

@@

Consider an incompressible, Stokes flow inside a domain $\Omega$ containing $M$ erodable bodies. We take the outer boundary $\Gamma$ to be a slightly smoothed version of the boundary of $[-3,3] \times [-1,1]$. The fluid flow is primarily from left to right, so that the inlet and outlet are located at approximately $x=-3$ and $+3$ respectively (the actual locations are slightly curved versions of these vertical lines due to domain smoothing). The erodable bodies, with boundaries $\gamma_\ell$, $\ell = 1,\ldots,M$, all sit inside of the central region $[-1,1] \times [-1,1]$. The boundary of the fluid domain is thus $\bd \Omega = \Gamma \cup \gamma_1 \cup \cdots \cup \gamma_M$. 
The empty space to the left and right of $[-1,1] \times [-1,1]$ serve as buffer regions to allow the flow profile imposed at the inlet and outlet to gradually adjust to the presence of the bodies. The equations governing the velocity $\uu$ and pressure $p$ of the fluid consist of the incompressible, Stokes equations coupled to boundary conditions:
\begin{equation}
\label{eqn:StokesEq}
  \begin{split}
    \mu \Delta \uu = \grad p, &\hspace{20pt} \xx \in \Omega, \gap 
      &&\mbox{\em conservation of momentum}, \\
    \grad \cdot \uu = 0, &\hspace{20pt} \xx \in \Omega, \gap 
      &&\mbox{\em conservation of mass}, \\
    \uu = \mathbf{0}, &\hspace{20pt} \xx \in \gamma, \gap 
      &&\mbox{\em no slip on the erodable bodies}, \\
    \uu = \UU, &\hspace{20pt} \xx \in \Gamma, \gap 
      &&\mbox{\em outer wall velocity}.
  \end{split}
\end{equation}
Above, $\UU$ represents the fluid velocity imposed along the outer boundary $\Gamma$, in particular at the inlet and outlet, as well as along the top and bottom walls. In this work, we impose a {\em uniform} flow profile along $\Gamma$, i.e.~$\UU = (U,0)$, although other choices are possible, for example a Poiseuille profile as employed in previous work \cite{quaife2018boundary, chiu2020viscous}. The advantages of the uniform profile are: (1) it will simplify the calculation of porous-medium properties, such as permeability and anisotropy, that will be described later; and (2), it may more realistically model the flow impinging upon a porous medium. We will allow the imposed flow speed to change with time $U = U(t)$ to enforce, for example, a desired pressure drop across the flow cell. To nondimensionalize the above system, we set the fluid viscosity to unity, $\mu = 1$.

@@

%-----------------------------------------------------------------------------------------------%
\section{Measuring porous-media properties: permeability, drag, and anisotropy}
\label{sec:}
%-----------------------------------------------------------------------------------------------%

@@

%-----------------------------------------------------------------------------------------------%
\section{Results}
%-----------------------------------------------------------------------------------------------%

@@

>>>>

\begin{abstract}
Using a Cauchy integral formulation of the boundary integral equations, we simulate the erosion a porous medium comprised of up to 100 solid bodies embedded in a Stokes flow. The grains of the medium are resolved individually and erode under the action of surface shear stress. Through nonlinear feedback with the surrounding flow fields, microscopic changes in grain morphology give way to larger-scale features in the medium such as channelization. The Cauchy-integral formulation and associated quadrature formulas enable us to resolve dense configurations of nearly contacting bodies.
We observe substantial anisotropy to develop over the course of erosion; that is, the configurations that result from erosion generally permit flow in the longitudinal direction more easily than in the transverse direction by up to a factor of five. These results suggest that the erosion of solid material from groundwater flows may contribute to the well-known anisotropy of natural porous media.
\end{abstract}
\maketitle

@@

	Figure \ref{fig1} shows an example simulation in which 80 circular bodies of randomized sizes and positions are immersed in a Stokes flow that moves from left to right. Color depicts the local speed of the flow intervening between bodies. Over time, individual bodies erode in response to the shear stresses induced on them, and heterogeneous material removal creates visible features both at the level of individual bodies and the larger-scale configuration. For example, horizontally-oriented channels are clearly visible in the third and fourth frames, and these contribute to the overall anisotropy of the medium.

% preferentially removing solid material, and further opening the channels. The positive feedback creates the channels that are clearly visible by in final two images.

@@

The outline of the paper is as follows. In Section \ref{sec:formulation} we discuss the governing equations for the fluid flow and for interface evolution. In Section \ref{sec:DLP} we discuss the numerical methods. In Section \ref{sec:medium} we measure porous-medium properties, in particular permeability.
In Section \ref{sec:results} we provide results and we conclude in section \ref{sec:conclusion}.

@@

Consider an incompressible, Stokes flow inside a domain $\Omega$ containing $M$ erodable bodies. We take the outer boundary $\Gamma$ to be a slightly smoothed version of the boundary of $[-3,3] \times [-1,1]$. The fluid flow is primarily from left to right, so that the inlet and outlet are located at approximately $x=-3$ and $+3$ respectively (the actual locations are slightly curved versions of these vertical lines due to domain smoothing). The erodable bodies, with boundaries $\gamma_\ell$, $\ell = 1,\ldots,M$, all sit inside of the central region $[-1,1] \times [-1,1]$. The boundary of the fluid domain is thus $\bd \Omega = \Gamma \cup \gamma_1 \cup \cdots \cup \gamma_M$. 
The empty space to the left and right of $[-1,1] \times [-1,1]$ serve as buffer regions to allow the flow profile imposed at the inlet and outlet to gradually adjust to the presence of the bodies. The equations governing the velocity $\uu$ and pressure $p$ of the fluid consist of the incompressible, Stokes equations coupled to boundary conditions:
\begin{equation}
\label{eqn:StokesEq}
  \begin{split}
    \mu \Delta \uu = \grad p, &\hspace{20pt} \xx \in \Omega, \gap 
      &&\mbox{\em conservation of momentum}, \\
    \grad \cdot \uu = 0, &\hspace{20pt} \xx \in \Omega, \gap 
      &&\mbox{\em conservation of mass}, \\
    \uu = \mathbf{0}, &\hspace{20pt} \xx \in \gamma, \gap 
      &&\mbox{\em no slip on the erodable bodies}, \\
    \uu = \UU, &\hspace{20pt} \xx \in \Gamma, \gap 
      &&\mbox{\em prescribed outer wall velocity}.
  \end{split}
\end{equation}
Above, $\UU$ represents the fluid velocity imposed along the outer boundary $\Gamma$, in particular at the inlet and outlet, as well as along the top and bottom walls. In this work, we impose a {\em uniform} flow profile along $\Gamma$, i.e.~$\UU = (U,0)$, although other choices are possible, for example a Poiseuille profile as employed in previous work \cite{quaife2018boundary, chiu2020viscous}. The advantages of the uniform profile are: (1) it will simplify the calculation of porous-medium properties, such as permeability and anisotropy, that will be described later; and (2), it may more realistically model the flow impinging upon a porous medium. We will allow the imposed flow speed to change with time $U = U(t)$ to enforce, for example, a desired pressure drop across the flow cell. To nondimensionalize the above system, we set the fluid viscosity to unity, $\mu = 1$.

@@

%-----------------------------------------------------------------------------------------------%
\section{Measuring porous-media properties: permeability, drag, and anisotropy}
\label{sec:medium}
%-----------------------------------------------------------------------------------------------%

@@

%-----------------------------------------------------------------------------------------------%
\section{Results}
\label{sec:results}
%-----------------------------------------------------------------------------------------------%

@@

%-----------------------------------------------------------------------------------------------%
\section{Conclusion}
\label{sec:conclusion}
%-----------------------------------------------------------------------------------------------%
